### 2018-07-18 paper/paper.tex

\documentclass[9pt,twocolumn,twoside,]{pnas-new}

%% Some pieces required from the pandoc template
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\templatetype{pnasresearcharticle}  % Choose template

\title{Expectations bias moral evaluations}

\author[a,1]{Derek Powell}
\author[b]{Zachary Horne}

  \affil[a]{Stanford University, Department of Psychology, 450 Serra Mall, Stanford,
CA, 94305}
  \affil[b]{Arizona State University, School of Social and Behavioral Sciences,
13591 N 47th Ave, Phoenix, AZ 85051}

% Please give the surname of the lead author for the running footer
\leadauthor{Powell}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{In two behavioral studies, we found that people's expectations biased
their reactions to morally-significant events--people were more upset by
surprising harms than more expected harms. This bias threatens to impose
a pernicious cycle: People living in poverty, unstable regions, and
other dangerous circumstances have the greatest need for assistance and
concern from the world at-large. Yet for this very reason, it is often
unsurprising when they are harmed. By blunting observers' concern for
the disadvantaged, this bias could make people less likely to donate
money, to volunteer, or to take political action, thereby increasing
inequality and compounding suffering. Recognizing and understanding this
bias will suggest ways to mitigate its harmful consequences.}

\authorcontributions{Both authors jointly conceived of the project and theory, designed and
conducted experiments, and analyzed results. D.P. developed the formal
aspects of the theory.}

\authordeclaration{The authors declare no conflicts of interest.}

\correspondingauthor{\textsuperscript{} }

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
 \keywords{  evaluation |  moral judgment |  information theory   }

\begin{abstract}
When we learn of tragic events--a violent killing or a terrorist
attack--we are mournful and often driven to action. But responses to
these events can differ substantially depending on the context in which
they occur. When tragedy takes us by surprise, we are shocked and
outraged, but when it strikes in unstable regions or rough
neighborhoods, we are often resigned to feel that ``these things
happen.'' Here we explain the role that expectations seem to play in
moral evaluation by modeling evaluations as the comparison of the state
of the world before and after an event. According to this theory,
expectations influence evaluations through information-theoretic
principles: less expected events do more to inform us about the state of
the world than do more expected events. Consistent with this theory, in
two preregistered studies we found that people's judgments of
morally-significant events were affected by the prior probability of
that event. People were more upset about an event that was unexpected
(e.g., a robbery at a clothing store) than an event that was more more
probable (e.g., a robbery at a convenience store). This bias threatens
to impose a pernicious cycle: People living in poverty, geo-politically
unstable regions, and other dangerous circumstances have the greatest
need for assistance and concern. However, for these very reasons, it is
often unsurprising when they are harmed. Consequently, the influence of
expectations on moral evaluations threatens to reduce observers' concern
for those most in need.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.

\acknow{We would like to acknowledge the help of Larisa Hussak, Keith Holyoak,
Alan Fiske, Hongjing Lu, John Hummel, Andrei Cimpian, and Ellen Markman
for their comments and support.}

When we learn of tragic events---a friend's loss of a loved-one, a
violent killing, or a terrorist attack---we are mournful, concerned,
angry, and driven to action. But our responses to these events differ
depending on the context in which they occur (1). In particular, our
expectations play an important role in our reactions to events. This
seems true of the ordinary: when a friend's loved-one passes, we often
ask, ``was it sudden?''; the dramatic: a murder in a sleepy town prompts
a very different reaction than a similar crime in a major city; and even
the catastrophic: acts of terrorism seem to evoke more outrage, horror,
and concern when they occur in peaceful rather than unstable regions.
For instance, the 2015 Paris attacks were felt around the world. Yet,
most of those mourning had little to say 15 hours earlier, when another
tragic attack killed at least 43 people in Beirut (2). Likewise, stories
of the Virginia Tech massacre, which killed 32 people, wildly
overshadowed coverage of attacks killing nearly 200 people in Iraq---one
of the bloodiest days since the 2003 invasion (3). In each of these
situations, the surprise elicited by an event appears to have affected
people's reactions to it. In some contexts, harm is tragic. In others,
``these things happen.''

Although this pattern seems so intuitive as to be obvious, no reasonable
code of ethics would justify these judgments, nor do any moral
psychological frameworks explicate the role of expectations in moral
evaluations. Outside of the domain of morality, everyday experiences and
controlled experiments show that people's evaluations of events
unambiguously depend on their expectations about those events. There is
a special thrill to having one's expectations exceeded and
disappointment when events fail to meet expectations: we root for the
underdog, hold surprise parties, and foreshadow bad news to ease its
delivery (4). Indeed, in a controlled study, Mellers and colleagues (5)
found that expectations influenced affective reactions during a gambling
task: Participants were often more excited by an small unexpected gain
(e.g., winning \$30 with 10\% probability) than they were by a larger
but more expected gain (e.g., winning \$60 with 90\% probability) (6).

It seems intuitive that surprise would shape people's reactions to
parties, sporting events, and games of chance. However, it seems wholly
inappropriate that surprise would shape people's reactions to events
with serious moral implications. Tragedies are tragic, whether expected
or not, and failure to recognize this could lead to tacit endorsement of
the pain and suffering of others wherever pain and suffering are the
status quo. If expectations truly do influence our reactions to
morally-significant world events, as intuition and anecdote seem to
suggest, there are potentially serious social consequences. To
demonstrate the plausibility of this suggestion, we develop a
theoretical account of how expectations influence evaluations, not as a
strange quirk or uniquely human bias, but instead as a fundamental
component of how people evaluate events.

Events, as we will conceive of them, are changes in the state of the
world---transitions from one state to another. For this reason, we
construe event evaluation as a comparison between the states of the
world before and after the event. Similar comparison processes play
crucial roles in many theories of decision-making (4, 5, 7, 8). For
instance, in Prospect Theory (9, 10), the utility of a prospect is
evaluated by comparison to a current reference point. Our proposal is
similar: we argue that expectations can set the reference points against
which people compare future outcomes, so that assigning value to an
event consists in comparing the state of the world following the event
to the state of the world just prior to the event. Further, we assume
that people's knowledge of the world is uncertain, so we represent the
states of the world probabilistically. This means that a person's model
for the state of the world is simply their expectations about what has
happened, what is happening, and what will happen in the future. A
direct implication of this proposal is that the evaluation of an event
is linked to what is learned about the world as a result of that event's
occurrence. That is, we evaluate events positively or negatively to the
extent that they inform our understanding of the world in positive or
negative ways.

One striking implication of this account is that it provides an
information-theoretic route to understanding how expectations directly
influence evaluation. A fundamental insight of Information Theory is
that the information carried by an event is a function of its prior
probability; low probability events carry more information than high
probability events (11). We propose that people learn more about the
state of the world when their expectations are violated by shocking
world events as compared to when they are affirmed by less surprising
events. For example, compared to a bombing in Lebanon, if a bombing
occurs in Paris, a less expected location, we learn that the world is
more dangerous than we had previously believed. Although this suggestion
is novel in the moral domain, these points are widely accepted
elsewhere. For instance, expectations have long been recognized as
fundamental to models of human and animal learning (12).

A more formal treatment of this theory establishes the potential for its
generality across domains. Let the state of the world prior to some
event be represented by a random variable, \(X\). We'll assume \(X\) is
discrete, and that every realization \(x\) represents some potential
specific state of the world in all relevant aspects under consideration.
An agent has some utility function \(u\), that applies over the
different realizations of \(X\). Thus the expected utility of the
present state of the world is:

\begin{equation*}
E_X[u(X)] = \sum_x u(x)p(x)
\end{equation*}

When we learn an event has occurred, we can represent this as an
observation of a second, binary random variable, \(S\). We can think of
\(S\) as a ``signalling event'' that informs us about about \(X\) (13).
That is, we assume there is a dependency between \(S\) and \(X\), so
that observing \(S\) informs us about the state of the world, \(X\). For
instance, reading a news report of a terror attack in Brussels informs
us about the state of the world, telling us about what realizations of
\(X\) are no longer possible (e.g., any wherein it was a peaceful day in
Brussels) and which are more or less likely. Conditioned on \(S\), the
expected utility of the world is:

\begin{equation*}
E_{X|S}[u(X)] = \sum_x u(x)p(x|s)
\end{equation*}

Our construal of event evaluation is as a comparison between states of
the world following and prior to an event. Consequently, we define the
value assigned to the signalling event, \(V(S)\), as:

\begin{equation*}
V(S) = E_{X|S}[u(X)] - E_X[u(X)]
\end{equation*}

From this it can be shown that the value assigned to event \(S\) is
proportional to its prior probability (see Theoretical Formalization).
Specifically, it is the difference in the expected utility under
different values of \(S\), weighted by the inverse of the prior
probability of \(S\).\footnote{It should be recognized that the result
  we obtain here bears a similarity to the ``disappointment function''
  included in Mellers and colleagues' (5) Decision Affect Theory.}

\begin{equation*}
V(s=1)=(1-p(s=1))\sum_x u(x)p(x|s=1) - u(x)p(x|s=0)
\end{equation*}

This conclusion does not depend on the nature of the event in question
nor on the utility function over events, but instead holds generally
wherever we conceive of evaluation as a holistic comparison between
uncertain states of the world. Consequently, this account predicts
expectations will equally affect reactions to gambles with clearly
specified values and probabilities (5), as well as real-world events
with ill-defined values, where reasoners' expectations are defined by
their existing beliefs and world knowledge.

Given the potentially fundamental role that expectations play in event
evaluation, and the potentially serious moral biases that may result, we
sought to test empirically whether people's evaluations of morally
harmful events are affected by their expectations about those events. We
conducted two preregistered studies examining the role of expectations
in moral evaluations. In both studies, participants were presented with
a series of trials in which they read brief descriptions of two
different events and were asked to indicate which of the two events
seemed more upsetting. In these trials, the two events were highly
similar, but were manipulated so that they differed in their perceived
prior probabilities: one event was more expected and one more
unexpected.

The design of these studies was chosen with special consideration to two
theoretical points. First, the information-theoretic account we have
described makes a solely directional prediction---all else being equal,
more surprising events will elicit stronger reactions than less
surprising events. Asking participants to compare two similar events
allowed us to test this directional prediction. Second, it is the
informativeness of events, determined by one's prior expectations of the
event, that we predicted would shape evaluations. Thus, reasoners should
not have to be induced to hold expectations experimentally (5), but
instead their own prior beliefs should be sufficient to shape their
expectations and affect their evaluations. For this reason, we
manipulated the prior probability of events by describing two otherwise
identical events occurring in different contexts. We explicitly chose
contexts that differed in ways that could not be accounted for by
construal-level explanations (1) or explanations based on biases towards
one's ingroup (14). This both allowed us to test the role of
expectations based on world knowledge and provided a more naturalistic
set of items.

\begin{figure}
\centering
\includegraphics{fig1.eps}
\caption{Probability with which participants chose the less expected
event as more upsetting across trials in Study 1. As predicted,
participants exhibited a robust bias toward choosing the less expected
event as more upsetting across all items. Error bars represent 95\%
credible intervals calculated from the hierarchical logistic regression
model. See Supporting Information for full description of items. {}}
\end{figure}

Consistent with our predictions, manipulating expectations shaped
evaluations across a range of morally-significant events. We found that
people viewed unexpected harmful events as more upsetting than expected
harmful events, even when the harm a victim suffered was the same. For
instance, people judged that a robbery at a clothing store (unexpected)
was more upsetting than a robbery at a convenience store (expected).
Furthermore, these differences were large, systematic, and robust.
Although there was variability in the situations that elicited the
largest differences, our statistical approach treated trial as a random
effect (see Methods and Results), allowing us to generalize our
conclusions to items we did not test (15).

\begin{figure}
\centering
\includegraphics{fig2.eps}
\caption{Participants' responses by item in Study 2. Upper: heatmap plot
displaying probability of choosing each response on the 1 to 5 scale
across items. Lower: marginal expected response across items, with 95\%
credible interval calculated from the ordinal regression model. Note the
scale has been constrained to show relevant region of results. In both
plots, a response of 3 indicates indifference, responses below 3
indicate a bias to choose the more expected event as more upsetting, and
responses above 3 indicate a bias to choose the less expected event as
more upsetting. Similar to Study 1, participants showed a robust bias
toward choosing the less expected event as more upsetting. Trials are
sorted by response to aid readability. {}}
\end{figure}

We have argued that expectations shape our reactions to events because
they determine how informative those events are in shaping our
understanding of the world. Thus, we argue that well-grounded
information theoretic principles can account for people's divergent
moral reactions to likely and unlikely events.

In some contexts, it is clearly adaptive that the violation of our
expectations would direct our attention to the events and situations
about which our understanding is most in need of revision (12). For
instance, an act of terrorism in a peaceful nation might be less
expected because of our beliefs about the effectiveness of security
forces, the stability of national relationships, and positive
relationships among people groups in that nation. Therefore, a
surprising act of terrorism casts doubt on our deeper understanding of
the geopolitical situation, and thereby leads us to suspect similar acts
are more likely to occur in the future.

Yet, however well-grounded the underlying principles, there is a clear
moral danger whenever expectations influence our moral concerns. In
particular, these effects pose a significant threat when considering our
reactions to humanitarian challenges brought on by poverty or regional
instability. In these cases, the influence of expectations on moral
evaluations threatens to impose vicious and morally pernicious cycles.
For instance, people living in geo-politically unstable regions or in
the developing world are often those who are most affected by terrorism,
famine, and natural disasters, and are the very people in greatest need
of assistance and concern from the world at-large. However, for
precisely these reasons, it is often unsurprising when harm befalls
people living in these circumstances. Thus the influence of expectations
on moral evaluations threatens to reduce observer's concern for these
victims most in need, making people less likely to donate time or money
to aid victims, or to take political action to prevent further harms.
The present research identifies this potentially pernicious moral bias
and grounds it in information theoretic principles. Hopefully, by
further understanding these cognitive dynamics, psychologists and
policymakers can begin to find ways to reduce or counteract people's
tendency to ignore the plight of those in most need of attention.

\matmethods{

We conducted two preregistered studies examining the role of
expectations in moral evaluations (\url{https://osf.io/86rsw/}). Each
study consisted of multiple trials where participants were presented
brief descriptions of two events. Their task was to decide which of the
two events was more upsetting. In Study 1, they made a two-alternative
forced-choice, and in Study 2 they made their judgments using a
five-point rating scale. Both studies consisted of 16 experimental
trials, five ``equivalent'' filler trials, and five ``non-equivalent''
filler trials.

In experimental trials, the two events described identical harms, but
one event was more expected and one more unexpected, as determined in a
prior norming study (see SI). For example, participants considered the
following stimulus:

\begin{itemize}
\item
  ``A 30 year old man in California dies in an earthquake''
  {[}Expected{]}
\item
  ``A 30 year old man in Oklahoma dies in an earthquake''
  {[}Unexpected{]}
\end{itemize}

In each of these events, the harm to the victim is the same (death) but
one event is more probable than the other, given the different
likelihoods of earthquakes occurring in California versus Oklahoma. In
addition to being more naturalistic, manipulating expectations through
(otherwise irrelevant) contextual details helps ensure that the harm
suffered by the victim is understood to be identical. In the context of
gambles, it is possible to manipulate probabilities specifically and
directly (5). However, in the context of more realistic moral harms, any
attempt to directly manipulate probabilities risks also changing
participants' perceptions of the degree of harm. For instance, suppose
we describe two cancer patients, one with a 5\% chance of survival and
one with a 50\% chance of survival. Even if they both ultimately meet
the same fate, participants might fairly assume the patient with the
lower chance of survival had a more serious illness, was more
debilitated, suffered more severely, and so forth. In contrast, changing
the context of the two events explains away these likelihood
differences, better equating the events in other respects.

Both studies also included ``equivalent'' filler trials. In these
trials, the two events differed in trivial contextual details that we
did not expect would affect participants' judgments. These filler trials
were meant to prevent participants from becoming explicitly aware of the
structure of the experimental trials. Finally, both studies included
``non-equivalent'' filler trials, the two events differed substantially
in the degree of harm suffered by a victim, so that one event was
expected to be seen as considerably more upsetting than the other. These
trials were included to allow participants a chance to use the extremes
of the response scale and to reduce any task demands that might drive
them to make artificially fine-grained distinctions when generating
their responses. All experimental materials for these studies are
available as Supporting Information at \url{https://osf.io/86rsw/}.

\subsection*{Exclusions}\label{exclusions}
\addcontentsline{toc}{subsection}{Exclusions}

In both studies, we excluded participants who failed to correctly answer
attention-check questions or who indicated at the end of the study they
had not taken their participation seriously.

\subsection*{Data Analysis}\label{data-analysis}
\addcontentsline{toc}{subsection}{Data Analysis}

We analyzed our data by performing Bayesian estimation using the
probabilistic programming language Stan (16). We tested our
preregistered predictions by computing Bayes Factors (i.e. \(BF_{01}\))
using the hypothesis function in the R package brms. Bayes Factors
express the ratio of the probability of data under the null hypothesis
to the probability of the data under an alternative hypothesis. Larger
Bayes Factors indicate that the data are more likely under the null
hypothesis (i.e., that the intercept is not different from zero) than
the alternative hypothesis (i.e., that the intercept is different from
zero), and vice versa. As Bayes Factors can be influenced by prior
choices (17), we performed prior robustness checks to confirm that our
conclusions would not depend substantially on the specification of the
priors.

To ensure the generality of our findings, we tested participants'
reactions to a variety of items and treated both participants and items
as random effects throughout our analyses. The use of random effects
over items allows us to consider these items as a sample from a larger
population of possible items, and to generalize our findings to that
wider population, much as we generalize from our sample of participants
to the larger population (15).

\subsection*{Study 1}\label{study1}
\addcontentsline{toc}{subsection}{Study 1}

\subsubsection*{Participants}\label{s1-participants}
\addcontentsline{toc}{subsubsection}{Participants}

A total of 255 participants were recruited from Amazon's Mechanical Turk
work distribution website (mTurk). Of these, 230 passed attention checks
and were included in the final analyses (105 male, 125 female, median
age = 33 years old). All participants were paid \$0.60 for their
participation.

\subsubsection*{Materials and procedure}\label{s1-materials-procedure}
\addcontentsline{toc}{subsubsection}{Materials and procedure}

Participants judged 16 experimental event-pairs and 10 equivalent filler
event-pairs. On each trial (an example is shown above), participants
were presented with the event-pair stimulus and had to judge which
outcome was more upsetting in a two-alternative forced choice task. The
two events were labeled ``Outcome 1'' and ``Outcome 2'' and the order,
and whether Outcome 1 or 2 was the unexpected event, was randomized and
counter-balanced.

\subsubsection*{Results and discussion}\label{s1-results}
\addcontentsline{toc}{subsubsection}{Results and discussion}

We predicted that people would judge that events where unexpected harm
occurred were more upsetting than events where expected harm occurred.
As indicated in Figure 1, we observed this trend. We estimated the
magnitude of this effect by fitting a Bayesian logistic random effects
model with participants' responses as the dependent variable (1 =
unexpected event is more upsetting; 0 = expected event is more
upsetting) and random intercepts for item and subject, allowing us to
generalize both to items and subjects we did not test. The intercept in
this model represents the log-odds of selecting the unexpected event as
being more upsetting. Thus, by examining the population-level intercept,
we can test whether participants were biased toward selecting the
unexpected event (\(\beta\) \textgreater{} 0), the expected event
(\(\beta\) \textless{} 0), or were unbiased (\(\beta\) = 0). We found
that people were considerably more likely to think that unexpected
events were more upsetting than events that were expected, Intercept =
1.424, 95\% CI {[}1.147, 1.707{]}, \(BF_{01}\) \textless{} .001. Figure
1 shows participants' responses broken-down by individual items.
Participants' bias toward selecting the unexpected event as more
upsetting was consistent across the 16 experimental items.

\subsection*{Study 2}\label{study2}
\addcontentsline{toc}{subsection}{Study 2}

In Study 1 we found that people's judgments of events were biased by
their expectations about those events. When forced to choose between two
events, participants decided that unexpected events were more upsetting
than expected events. In Study 2, we sought to test our hypothesis using
a more conservative method: we provided participants with a more
expressive response scale so that if they viewed the events under
consideration as equally harmful, their responses could reflect their
attitude.

\subsubsection*{Participants}\label{s2-participants}
\addcontentsline{toc}{subsubsection}{Participants}

A total of 504 participants were recruited from Amazon's Mechanical Turk
work distribution website (mTurk). Of these, 372 passed attention checks
and were included in the final analyses (174 male, 198 female, median
age = 34 years old). All participants were paid \$0.60 for their
participation.

\subsubsection*{Materials and procedure}\label{s2-materials-procedure}
\addcontentsline{toc}{subsubsection}{Materials and procedure}

The materials in Study 2 were the same as those in Study 1. As in Study
1, on each trial of the study, participants were presented with a pair
of actions labeled ``Outcome 1'' and ``Outcome 2'' and were asked,
``Which outcome seems more upsetting?'' In Study 2 participants made
their rating on a five-point scale (Outcome 1 seems more upsetting,
Outcome 1 seems a little more upsetting, neither seems more upsetting
than the other, Outcome 2 seems a little more upsetting, Outcome 2 seems
more upsetting).

\subsubsection*{Results and discussion}\label{s2-results}
\addcontentsline{toc}{subsubsection}{Results and discussion}

We predicted that participants would evaluate unexpected events as more
upsetting than expected events. We estimated the magnitude of this
effect by fitting a cumulative (ordinal) random effects model with
participants' scale responses as the dependent variable (1 to 5) and
random intercepts for item and subject. This model produces four
intercept coefficients, representing the cumulative log-odds of
responses at each scale point or lower. For instance, the second
coefficient represents the log-odds participants chose a 2 (``outcome 2
seems slightly more upsetting'') or lower on the scale. Similarly, the
third intercept coefficient represents the log-odds participants chose a
3 or lower on the scale. To examine the predicted effect, we compared
the third intercept coefficient (representing the log-odds of choosing 4
or higher) to the inverse of the second intercept coefficient (thereby
representing the log-odds \emph{not} choosing a 3 or higher---i.e.,
choosing a 1 or 2), allowing us to test whether participants were more
likely to choose the expected or unexpected event as being more
upsetting in cases where they did not choose the neither option. This
analysis indicated that people were more likely to think that events
that were unexpected were more upsetting than events that were expected,
\(BF_{01}\) \textless{} .001 (see supporting information for full model
results). We observed this difference consistently across nearly every
item (Figure 2).

We note that the most common response was to view both events, which
involve identical harms, as equivalently upsetting. However, across
almost every item, participants showed a clear bias towards choosing
unexpected events as more upsetting. Thus, Study 2 replicates the
tendency to view unexpected moral events as more upsetting using a more
conservative judgment task.

\section*{Theoretical Formalization}\label{theory}
\addcontentsline{toc}{section}{Theoretical Formalization}

We construe event evaluation as a comparison between the state of
affairs prior to an event and the state of affairs following the event.
We represent the state of the world as a random variable, \(X\). We
assume \(X\) is discrete, and that every realization \(x\) represents
some potential specific state of the world in all relevant aspects under
consideration. An agent has some utility function \(u\), that applies
over the different realizations of \(X\). Thus the expected utility of
the present state of the world is:

\begin{equation*}
E_X[u(X)] = \sum_x u(x)p(x)
\end{equation*}

When we learn an event has occurred, we can represent this as an
observation of another random variable, \(S\). We can think of \(S\) as
a ``signalling event'' because it informs (signals) to us something
about \(X\) (13). For simplicity, we treat \(S\) as binary, either 1 or
0. Importantly, we assume there is a dependency between \(S\) and \(X\),
so that observing \(S\) informs our knowledge of \(X\). For instance,
reading a news report of a terror attack in Brussels informs us about
the state of the world, telling us about what realizations of X are no
longer possible (e.g., any wherein it was a peaceful day in Brussels)
and which are more or less likely. Conditioned on \(S\), the expected
utility of the world is:

\begin{equation*}
E_{X|S}[u(X)] = \sum_x u(x)p(x|s)
\end{equation*}

We define the value assigned to the signalling event, \(V(S)\), as:

\begin{equation}
V(S) = E_{X|S}[u(X)] - E_X[u(X)]
\end{equation}

Which we write more specifically as:

\begin{equation}
V(s=1) = \sum_x u(x)p(x|s=1) - \sum_x u(x)p(x)
\end{equation}

Our goal is to express this relation in terms of \(p(S)\). First, we
manipulate \(E_X[u(X)]\) by applying the expectation of conditional
expectation and some algebra.

\begin{align*}
\sum_x u(x)p(x) & = \sum_s u(x)p(x|s)p(s) \\
 &= \begin{aligned}[t]
    & \sum_x u(x)p(x|s=1)p(s=1) - \\
    & \quad \sum_x u(x)p(x|s=0)p(s=0) \end{aligned}\\
 &= \begin{aligned}[t]
    & p(s=1)\sum_x u(x)p(x|s=1) -\\
    & \quad p(s=0) \sum_x u(x)p(x|s=0) \end{aligned}\\
 &= \begin{aligned}[t]
    & p(s=1)\sum_x u(x)p(x|s=1) - \\
    & \quad (1-p(s=1)) \sum_x u(x)p(x|s=0) \end{aligned}\\
\end{align*}

Substituting this quantity back into Equation 2, we obtain:

\begin{align*}
V(s=1) = & \sum_x u(x)p(x|s=1) - \\ & p(s=1)\sum_x u(x)p(x|s=1) -\\ & (1-p(s=1)) \sum_x u(x)p(x|s=0)
\end{align*}

Finally, we factor the first two sums and combine them with the last sum
to yield:

\begin{equation}
V(s=1)=(1-p(s=1))\sum_x u(x)p(x|s=1) - u(x)p(x|s=0)
\end{equation}

Under this formalization, the role of expectations in the evaluation of
events does not depend on any further assumptions about the states of
the world or the utility function to be applied over them. Instead, this
conclusion depends on just two assumptions: first, that events are
evaluated by comparing the utilities of the states of the world prior to
and following the event; and second, that knowledge of those states of
the world is uncertain.

We have appealed to the informativeness of an event to illustrate
intuitively how expectations should influence evaluations. More expected
events carry less information, and thereby engender weaker reactions.
Though our model of event evaluation is not an information-theoretic
model in a direct sense, this interpretation still applies: the
information carried by observing \(S\) is a function of that
observation's prior probability, which we have in turn shown determines
its evaluation.
}
\showmatmethods
\showacknow
\pnasbreak

\hypertarget{refs}{}
\hypertarget{ref-Trope2010}{}
1. Trope Y, Liberman N (2010) Construal-level theory of psychological
distance. \emph{Psychological review} 117(2):440--63.

\hypertarget{ref-Graham2015}{}
2. Graham DA (2015) The empathy gap between paris and beirut.

\hypertarget{ref-Greenslade2007}{}
3. Greenslade R (2007) A hierarchy of death.

\hypertarget{ref-Bell1985}{}
4. Bell D (1985) Disappointment In Decision Making Under Uncertainty.
\emph{Operations Research} 33(1):1--27.

\hypertarget{ref-Mellers1997}{}
5. Mellers B a, Schwartz a, Ho K, Ritov I (1997) Decision Affect Theory:
Emotional Reactions to the Outcomes of Risky Options.
\emph{Psychological Science} 8(6):423--429.

\hypertarget{ref-Shepperd2002}{}
6. Shepperd J a, Mcnulty JK (2002) The affective consequences of
expected and unexpected outcomes. \emph{Psychological science}
13(1):85--88.

\hypertarget{ref-Gul1991}{}
7. Gul F (1991) A theory of disappointment aversion. \emph{Econometrica}
59(3):667--686.

\hypertarget{ref-Loomes1986}{}
8. Loomes G, Sugden R (1986) Disappointment and Dynamic Consistency in
Choice under Uncertainty. \emph{Review of Economic Studies}
53(2):271--282.

\hypertarget{ref-Kahneman1979}{}
9. Kahneman D, Tversky A (1979) Prospect Theory: An Analysis of Decision
under Risk. \emph{Econometrica} 47(2):263--292.

\hypertarget{ref-Tversky1992}{}
10. Tversky A, Kahneman D (1992) Advances in prospect theory: Cumulative
representation of uncertainty. \emph{Journal of Risk and uncertainty}
5:297--323.

\hypertarget{ref-Shannon1948}{}
11. Shannon CE (1948) A Mathematical Theory of Communication. \emph{Bell
System Technical Journal} 27(3):379--423.

\hypertarget{ref-Rescorla1972}{}
12. Rescorla RA, Wagner AR (1972) A theory of Pavlovian conditioning:
Variations in the effectiveness of reinforcement and nonreinforcement.
\emph{Classical Conditioning II Current Research and Theory}, pp 64--99.

\hypertarget{ref-Arrow1996}{}
13. Arrow KJ (1996) The economics of information: An exposition.
\emph{Empirica} 23(2):119--128.

\hypertarget{ref-Brewer1999}{}
14. Brewer MB (1999) The psychology of prejudice: Ingroup love or
outgroup hate? \emph{Journal of Social Issues} 55(3):429--444.

\hypertarget{ref-Judd2012}{}
15. Judd CM, Westfall J, Kenny DA (2012) Treating stimuli as a random
factor in social psychology: A new and comprehensive solution to a
pervasive but largely ignored problem. \emph{Journal of Personality and
Social Psychology} 103(1):54--69.

\hypertarget{ref-Carpenter2017}{}
16. Carpenter B, et al. (2017) Stan : A Probabilistic Programming
Language. \emph{Journal of Statistical Software} 76(1).
doi:\href{https://doi.org/10.18637/jss.v076.i01}{10.18637/jss.v076.i01}.

\hypertarget{ref-Gelman2017}{}
17. Gelman A, Simpson D, Betancourt M (2017) The Prior Can Often Only Be
Understood in the Context of the Likelihood. \emph{Entropy} 19(10):555.

% Bibliography
% \bibliography{pnas-sample}

\end{document}