### 2018-09-06 SuppInfo.Rmd
paper edits pre submission

---
title: "Expectations bias moral evaluations: Supporting information"
output:
  pdf_document:
bibliography: "paper/wiw-biblio.bib"
csl: apa6.csl
    # keep_tex: true
  # html_notebook: default
---

```{r, include=FALSE}
library(tidyverse)

df_raw <- read.csv("data/WIW_Norming_for_preregistered_study.csv")

reverse_items <- c(
  "Earthquake",
  "Tornado",
  "Fastfood",
  "Marine",
  "Police",
  "Mugging",
  "Crushed",
  "Dysentery",
  "Drunk",
  "Football"
)

df_wide <- df_raw %>%
  filter(Status==0) %>%
  filter(Check_trial == 5) %>%
  select(-Check_trial) %>%
  mutate(sex = if_else(sex == 1, "male", "female"))

df <- df_wide %>%
  gather(trial, response, Earthquake:Rape) %>%
  mutate(response = ifelse(trial %in% reverse_items, 6-response, response))
```

# Theoretical Formalization

We construe event evaluation as a comparison between the state of affairs prior to an event and the state of affairs following the event. We represent the state of the world as a random variable, $X$. We assume $X$ is discrete, and that every realization $x$ represents some potential specific state of the world in all relevant aspects under consideration. An agent has some utility function $u$, that applies over the different realizations of $X$. Thus the expected utility of the present state of the world is:

\begin{equation*}
E_X[u(X)] = \sum_x u(x)p(x)
\end{equation*}

When we learn an event has occurred, we can represent this as an observation of another random variable, $S$. We can think of $S$ as a "signalling event" because it informs (signals) to us something about $X$ [this formalism follows @Arrow1996]. For simplicity, we treat $S$ as binary, either 1 or 0. Importantly, we assume there is a dependency between $S$ and $X$, so that observing $S$ informs our knowledge of $X$. For instance, reading a news report of a terror attack in Brussels informs us about the state of the world, telling us about what realizations of X are no longer possible (e.g., any wherein it was a peaceful day in Brussels) and which are more or less likely. Conditioned on $S$, the expected utility of the world is:

\begin{equation*}
E_{X|S}[u(X)] = \sum_x u(x)p(x|s)
\end{equation*}

We define the value assigned to the signalling event, $V(S)$, as:

\begin{equation}
V(S) = E_{X|S}[u(X)] - E_X[u(X)]
\end{equation}

Which we write more specifically as:

\begin{equation}
V(s=1) = \sum_x u(x)p(x|s=1) - \sum_x u(x)p(x)
\end{equation}

Our goal is to express this relation in terms of $p(S)$. First, we manipulate $E_X[u(X)]$ by applying the expectation of conditional expectation and some algebra.

\begin{align*}
\sum_x u(x)p(x) & = \sum_s u(x)p(x|s)p(s) \\
 &= \begin{aligned}[t]
    & \sum_x u(x)p(x|s=1)p(s=1) - \\
    & \quad \sum_x u(x)p(x|s=0)p(s=0) \end{aligned}\\
 &= \begin{aligned}[t]
    & p(s=1)\sum_x u(x)p(x|s=1) -\\
    & \quad p(s=0) \sum_x u(x)p(x|s=0) \end{aligned}\\
 &= \begin{aligned}[t]
    & p(s=1)\sum_x u(x)p(x|s=1) - \\
    & \quad (1-p(s=1)) \sum_x u(x)p(x|s=0) \end{aligned}\\
\end{align*}

Substituting this quantity back into Equation 2, we obtain:

\begin{align*}
V(s=1) = & \sum_x u(x)p(x|s=1) - \\ & p(s=1)\sum_x u(x)p(x|s=1) -\\ & (1-p(s=1)) \sum_x u(x)p(x|s=0)
\end{align*}

Finally, we factor the first two sums and combine them with the last sum to yield:

\begin{equation}
V(s=1)=(1-p(s=1))\sum_x u(x)p(x|s=1) - u(x)p(x|s=0)
\end{equation}

Under this formalization, the role of expectations in the evaluation of events does not depend on any further assumptions about the states of the world or the utility function to be applied over them. Instead, this conclusion depends on just two assumptions: first, that events are evaluated by comparing the utilities of the states of the world prior to and following the event; and second, that knowledge of those states of the world is uncertain.

We have appealed to the informativeness of an event to illustrate intuitively how expectations should influence evaluations. More expected events carry less information, and thereby engender weaker reactions. Though our model of event evaluation is not an information-theoretic model in a direct sense, this interpretation still applies: the information carried by observing $S$ is a function of that observation’s prior probability, which we have in turn shown determines its evaluation.

\newpage

# Item Norming

We created and normed 20 items, each consisting of a pair of events that differed in their prior likelihood. The two events (outcomes) were manipulated so what one would seem more expected than the other. Participants (_N_ = `r nrow(df_wide)`) were asked to choose which event seemed more likely using a 5-point scale (1 = Event 1 seems a lot more likely, 2 = Event 1 seems a little more likely, 3 = Neither event seems more likely, 4 = Event 2 seems a little more likely, 5 = Event 2 seems much more likely). As predicted, for all items the "more expected" outcome was judged to be more likely. Plot below shows mean likelihood rating by item along with approximate 95% confidence intervals.

```{r, echo=FALSE}
# standard error function
stderr <- function(x) {
          sqrt(var(x[!is.na(x)]) / length(x[!is.na(x)]))
}

df %>%
  group_by(trial) %>%
  summarize(
    Mean = mean(response),
    UL = mean(response) + 1.96*stderr(response),
    LL = mean(response) - 1.96*stderr(response)
    ) %>%
  ggplot() +
  aes(x = reorder(trial, -Mean), y = Mean, ymin = LL, ymax=UL) +
  geom_pointrange() +
  geom_hline(yintercept = 3, linetype="dashed") +
  coord_cartesian(ylim=c(1,5)) +
  theme_bw()+
  labs(x = "Trial", y="Mean Likelihoods Judgment") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

From among these 20 items we selected 16 items for use in our experiments.

\newpage
## Items

Below we provide the full text of items as presented to participants.

### Experimental Trials

| Trial      | More Expected                                                | Less Expected                                                |
| ---------- | ------------------------------------------------------------ | ------------------------------------------------------------ |
| Tornado    | A family in Nebraska has their home destroyed by a tornado.  | A family in Oregon has their home destroyed by a tornado.    |
| Marine     | A young US marine is shot in the head and killed by friendly fire on patrol outside Baghdad. | A young US marine is shot in the head and killed by friendly fire during a training exercise. |
| Mugging    | A man in downtown Detroit is mugged.                         | A man in downtown Denver is mugged.                          |
| Crushed    | A worker is crushed to death due to an accident at a factory in India. | A worker is crushed to death due to an accident at a factory in Japan. |
| Dysentery  | A pair of travelers contract dysentery after drinking water in Mexico. | A pair of travelers contract dysentery after drinking water in Arizona. |
| Cancer     | A 60 year old woman dies from breast cancer.                 | A 60 year old man dies from breast cancer.                   |
| Drunk      | A pedestrian is killed by a drunk driver on a Friday night.  | A pedestrian is killed by a drunk driver on a Saturday morning. |
| Football   | A 16 year old sustains a major concussion while playing in a high school football game. | A 16 year old sustains a major concussion while playing in a high school basketball game. |
| Heater     | A woman on disability dies when her 15 year-old heater fails during winter. | A woman on disability dies when her 3 year-old heater fails during winter. |
| Flattire   | A man repairing a flat tire on the side of the highway is struck by a car going 50 mph. | A man repairing a flat tire on the side of a suburban street is struck by a car going 50 mph. |
| Clerk      | A clerk is held up at gunpoint in a convenience store.       | A clerk is held up at gunpoint in a clothing store.          |
| Corruption | A corrupt government official in Mexico takes a bribe that allows a company to pollute the environment. | A corrupt government official in Canada takes a bribe that allows a company to pollute the environment. |
| Rape       | A woman is raped leaving a bar.                              | A woman is raped leaving a coffee shop.                      |
| Earthquake | A 30 year old man in California dies in an earthquake.       | A 30 year old man in Oklahoma dies in an earthquake.         |
| Fastfood   | A 32 year old woman gets food poisoning after eating a hamburger at a fast food restaurant. | A 32 year old woman gets food poisoning after eating a hamburger at a four star restaurant. |
| Police     | A criminal shoots and wounds a police officer while the officer is responding to a 911 call reporting a robbery in progress. | A criminal shoots and wounds a police officer while the officer is responding to a 911 call reporting vandalism. |

\newpage

### Filler Trials

| Trial    | Outcome 1                                                    | Outcome 2                                                    |
| -------- | ------------------------------------------------------------ | ------------------------------------------------------------ |
| Check1   | This item tests if you are paying attention                  | Please choose event 2 seems a little more upsetting          |
| Check2   | Someone who is paying attention will select event two        | Please select event two seems a lot more upsetting.          |
| Check3   | A participant who is paying attention will select event one  | Please select event one seems a little more upsetting.       |
| NoDiff1  | A man in Connecticut dies in a house fire.                   | A man in New Hampshire dies in a house fire.                 |
| NoDiff2  | A family in Florida is evicted from their home after they run into financial troubles. | A family in Louisiana is evicted from their home after they run into financial troubles. |
| NoDiff3  | After an accident, a machinist has to have his foot amputated. | After an accident, a metalworker has to have his foot amputated. |
| NoDiff4  | A girl skins her knees skateboarding.                        | A girl skins her knees rollerblading.                        |
| NoDiff5  | A large tree branch falls and injures a couple walking home from a baby shower. | A large tree branch falls and injures a couple walking home from their church. |
| BigDiff1 | A cat hurts its leg jumping off the refrigerator.            | A cat dies after eating spoiled food.                        |
| BigDiff2 | A parked car is damaged in a parking lot in Oregon.          | A car is totaled on the highway in Washington.               |
| BigDiff3* | \$1000 worth of products from an electronics store are destroyed in a shipping accident. | \$200 worth of products from a clothing store are burned in a shipping accident. |
| BigDiff4 | A knife is found in a high school student's locker.          | A gun is found in a high school student's gym bag.           |
| BigDiff5 | A 30 year-old man is killed in a car crash.                  | A 35 year-old man is injured in a cycling accident.          |

_* As presented in Study 2. In Study 1, this concerned a construction worker having his left versus right foot amputated, but participants judged having one's right foot amputated was somewhat worse than having one's left foot amptuated._

\newpage

## Discarded items

These items were included in the norming study but were not ultimately selected for the final study. "Heroin" and "theft" showed the weakest effects on likelihood judgments. We decided that the damage to property in "Scratched" was not sufficiently moral in nature. Finally, "brokeninto" was discarded for being too similar to "flattire" which we ultimately decided had more significant moral content.

| Trial      | Event 1                                                      | Event 2                                                      |
| ---------- | ------------------------------------------------------------ | ------------------------------------------------------------ |
| heroin     | A suburban teenage boy dies of a heroin overdose.            | A homeless teenage boy dies of a heroin overdose.            |
| theft      | A man in training steals \$50,000 from a small business he recently began working for. | A man on probation steals $50,000 from a small business he recently began working for. |
| scratched  | A woman's car is scratched in a quiet parking lot.           | A woman's car is scratched in a busy parking lot.            |
| brokeninto | Three cars on the same street are broken into while parked in the suburbs. | Three cars on the same street are broken into while parked downtown. |

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
\noindent